---
title: "Reproducible Research Week2 Coursera Assignment"
author: "Sakthi"
date: "January 7, 2017"
output: html_document
---

## R Markdown

Loading and preprocessing the data

1) Set Working directory
2) Read in data file
3) Process / transform the data


```{r setup, echo=TRUE}
setwd('C:/Sakthi/Study Materials/Data Science/Course - 5')
Inptfl <- read.csv('activity.csv')
Inptfl$date <- strptime(Inptfl$date, "%Y-%m-%d")
```

What is mean total number of steps taken per day?

1) Calculate the total number of steps taken per day
2) Make a histogram of the total number of steps taken each day
3) Calculate and report the mean and median of the total number of steps taken per day

```{r}

Inptfl1 <- Inptfl[c("steps", "date")]
Inptfl1 <- na.omit(Inptfl1)

# Aggregate based on date
library(plyr)
totalStepsPerDay <- ddply(Inptfl1, .(date), summarize, sum_x=sum(steps))

# Plot a (pseudo) histogram where the x-axis denotes the day and the y-axis denotes the total number of steps taken for each day :
plot(totalStepsPerDay, main="Histogram of steps taken each day", xlab="Date (October to November 2012)", ylab="Frequency", type="h", lwd=4, col="blue")

totalStepsPerDay <- ddply(Inptfl1, .(date), summarise, "sum"=sum(steps) ,"mean"=mean(steps)) 
#Median
median_x <- median(totalStepsPerDay$sum)
median_x

# Mean
mean_x <- mean(totalStepsPerDay$sum)
mean_x
```

What is the average daily activity pattern?

1) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}

# Split up the data according to the interval
intervalSplit <- split(Inptfl$steps, Inptfl$interval)

# Find the average amount of steps per time interval - ignore NA values
averageStepsPerInterval <- sapply(intervalSplit, mean, na.rm=TRUE)

# Keep a list of all possible intervals
uniqueIntervals <- unique(Inptfl$interval)

# Plot the time-series graph
plot(uniqueIntervals, averageStepsPerInterval, type="l",
     main="Average number of steps per interval across all days", 
     xlab="Interval", ylab="Average # of steps across all days", 
     lwd=2, col="blue")

# Find the location of where the maximum is
maxIntervalDays <- max(averageStepsPerInterval, na.rm=TRUE)
maxIndex <- as.numeric(which(averageStepsPerInterval == maxIntervalDays))

#Maximum number of steps
maxInterval <- uniqueIntervals[maxIndex]
maxInterval
```

Inputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
#Total number of missing values
numMissingValues <- sum(is.na(Inptfl$steps))
numMissingValues

#Fill missing values
library(Hmisc)
InptflImputed <- Inptfl
InptflImputed$steps <- impute(Inptfl$steps, fun=mean)

stepsByDayImputed <- ddply(InptflImputed, .(date), summarise, "steps"=sum(steps))

#Make Histogram
plot(stepsByDayImputed, main="Histogram of steps taken each day(Imputed)", xlab="Total steps per day (Imputed)", ylab="Frequency", type="h", lwd=4, col="blue")

#Median after Impute
median_x <- median(stepsByDayImputed$steps)
median_x

# Mean after Impute
mean_x <- mean(stepsByDayImputed$steps)
mean_x
```

Are there differences in activity patterns between weekdays and weekends?

1) Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
2) Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
InptflImputed$dateType <-  ifelse(as.POSIXlt(InptflImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')

averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=InptflImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```